---
title: "Penguins size comparisons"
author: "Nichoals Bissonnette"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests

mass comparison : male and female chinstrap penguis

```{r}
chinstraps <-penguins %>% 
  filter(species =="Chinstrap")

# Histograms 
ggplot(data=chinstraps, aes(x=body_mass_g)) +
  geom_histogram(bins = 10)+
  facet_wrap(~sex)

## QQ plot 
ggplot(data = chinstraps, aes(sample = body_mass_g))+
  geom_qq() +
  facet_wrap(~sex)

## get descriptive statistics 
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass= mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass= sd(body_mass_g),
    sample_size = n()
  )
```

get vectors of body_mass_g for female and male chinstrap penguins separtely.

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex== "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex== "male") %>% 
  pull(body_mass_g)
```

### T-tests a couple ways

1. Is ther a significant differce in body mass betwee nmale and female chinstrap penguins?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)
```

Are male chinstrap penhiuns significantly larger (body mass) than female chinstrap penguins? 

```{r}
t.test(chinstraps_m,chinstraps_f, alternative = "greater")
```

## one-way ANOVA

## 1.  table of counts by species and sex
```{r}
penguins_counts <- penguins %>% 
  count(species,sex)
penguins_counts
```

create a subset of only female penguins 
```{r}
penguins_f <- penguins %>% 
filter(sex== "female")

ggplot(data= penguins_f, aes(x= flipper_length_mm)) +
  geom_histogram(aes(fill=species))+
  facet_wrap(~species, scales= "free")

ggplot(data= penguins_f, aes(sample= flipper_length_mm)) +
  geom_qq(aes(color=species))+
  facet_wrap(~species, scales= "free")
```

### summary statistics 
Starting from penguins f, crete a summary tabe athe groups by specis, and returens the following:
-mean of lipper length
-sd""
-sample size
-variance'var'
-standard error= sd/squrt(n)

```{r}
penguin_f_summary <-penguins_f %>% 
  group_by(species) %>% 
  summarize(
    mean_flipper_length= mean(flipper_length_mm),
    sd_flipper_length =sd(flipper_length_mm),
    sample_size= n(),
    se_flipper_length = sd(flipper_length_mm)/sqrt(n()),
    var_flipper_length = var(flipper_length_mm)
    
  )

leveneTest(flipper_length_mm ~ species, data= penguins_f)
```
### Beeswarm plot with mean +- 1sd

```{r}
ggplot() +
  geom_beeswarm(data=penguins_f, 
                aes (x=species, y=flipper_length_mm),
                color= "gray70",
                size=3,
                alpha=.5) +
  geom_point(data= penguin_f_summary, aes(x=species, y= mean_flipper_length),
             size=4)+
  geom_errorbar(data= penguin_f_summary,
                aes(x=species,
                    ymax=mean_flipper_length +sd_flipper_length,
                    ymin= mean_flipper_length - sd_flipper_length),
                width=.1) +
  theme_light()
```

## ONe-way ANOVA

```{r}
penguin_aov <- aov(flipper_length_mm ~species, data= penguins_f)
summary(penguin_aov)

TukeyHSD(penguin_aov)
```

Type III Anova with the car package:
```{r}
penguin_lm <- lm(flipper_length_mm~ species, data= penguins_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
```

